---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ZooGVT

<!-- badges: start -->
<!-- badges: end -->

The goal of ZooGVT is to visualize and understand what genes give rise or contribute to what anatomical features. The user is able to see what adjective they use as a feature for a animals in the BgeeDB database that have phenotype descriptions in Ensemble.


## Description

ZooGVT allows users to visualize the various anatomical features related to a set of genes. This gives a better map of how certain genes translate into the animal features. These Genes can alter be extracted to compare differences between species. THis package allows easier access to the Uberon Ontolgy and aids in its use.


## Installation


The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Tsmnbx/ZooGVT")
```
## Example

This is a basic example which shows the UBERONids on could get for a Zebra fish with the descriptor "pectoral fin":

```{r example}
library(ZooGVT)
#BiocManager::install("topGO")

IdDataFrame<-ZooGVT::GetUberonIds("Danio_rerio","Zebrafish","pectoral fin")
IdDataFrame
```
To run the Shiny app
```{r eval=FALSE}
runZooGVT()
```


Here is the package structure
```{r}
print("├── DESCRIPTION
├── Danio_rerio_Bgee_14_1
│   ├── topAnat_AnatEntitiesNames_7955.tsv
│   ├── topAnat_AnatEntitiesRelationships_7955.tsv
│   └── topAnat_GeneToAnatEntities_7955_PRESENCESILVER.tsv
├── LICENSE
├── LICENSE.md
├── NAMESPACE
├── R
│   ├── GenesFromAnatomy.R
│   ├── GetUberonIds.R
│   ├── ListSpecies.R
│   └── data.R
├── README.Rmd
├── README.md
├── ZooGVT.Rproj
├── data
│   └── phyenotypeDescriptions.rda
├── inst
│   └── shiny-scripts
│       └── app.R
├── man
│   ├── GenesFromAnatomy.Rd
│   ├── GetUberonIds.Rd
│   ├── ListSpecies.Rd
│   ├── figures
│   │   └── README-pressure-1.png
│   └── phyenotypeDescriptions.Rd
├── release.tsv
├── species_Bgee_14_1.tsv
├── tests
│   ├── testthat
│   │   ├── release.tsv.tmp
│   │   └── test-ListSpecies.R
│   └── testthat.R
└── vignettes
    └── Introduction_ZooGVT.Rmd"
)
```





## Contributions

The *ListSpecies* Function uses the BgeeDB package. The *GetUberonIds* function also made use of the BgeeDB package which uses topGO as well as the biomaRt for access to Ensembl. *GenesFromAnatomy* used all of these plus fmsb to produce a radar graph.




## References

Komljenovic A, Roux J, Robinson-Rechavi M, and Bastian F (2016). BgeeDB,
an R package for retrieval of curated expression datasets and for gene
list enrichment tests. F1000Research

Bastian F et al. (2008). Bgee: Integrating and Comparing Heterogeneous
Transcriptome Data Among Species. Data Integration Life Sci. Lecture Notes
in Computer Science, 5109, 124-31

Charif, D. and Lobry, J.R. (2007)

Martin Morgan (2019). BiocManager: Access the Bioconductor Project Package
Repository. R package version 1.30.10.
https://CRAN.R-project.org/package=BiocManager

Mapping identifiers for the integration of genomic datasets with the
  R/Bioconductor package biomaRt. Steffen Durinck, Paul T. Spellman, Ewan
  Birney and Wolfgang Huber, Nature Protocols 4, 1184-1191 (2009).

BioMart and Bioconductor: a powerful link between biological databases
and microarray data analysis. Steffen Durinck, Yves Moreau, Arek
Kasprzyk, Sean Davis, Bart De Moor, Alvis Brazma and Wolfgang Huber,
Bioinformatics 21, 3439-3440 (2005).
  
  
Minato Nakazawa (2019). fmsb: Functions for Medical Statistics Book
  with some Demographic Data. R package version 0.7.0.
  https://CRAN.R-project.org/package=fmsb
  
Adrian Alexa and Jorg Rahnenfuhrer (2020). topGO: Enrichment Analysis for Gene Ontology. R package
  version 2.42.0.

R Core Team (2020). R: A language and environment for statistical computing. R Foundation for
  Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.  

“Learn Shiny.” Shiny, shiny.rstudio.com/tutorial/.

Holtz, Yan. “Basic Radar Chart.” – The R Graph Gallery, www.r-graph-gallery.com/142-basic-radar-chart.html.

Andrea Komljenovic, Julien Roux. BgeeDB, an R Package for Retrieval of Curated Expression Datasets and for Gene List Enrichment Tests, 27 Oct. 2020, bioconductor.org/packages/release/bioc/vignettes/BgeeDB/inst/doc/BgeeDB_Manual.html.

## Acknowledgements


This package was developed as part of an assessment for 2020 BCB410H: Applied Bioinformatics, University of Toronto, Toronto,CANADA.
