---
title: "An to Introduction ZooGVT"
author: "Joshua Efe"
output: rmarkdown::html_vignette 
vignette: >
  %\VignetteIndexEntry{A tour of ZooGVT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

##Introduction

ZooGVT allows a user to see what Phenotype descriptions have how many genes associated with them from the DgeeDb database

Use the following to download

```{r setup}
require("devtools")
devtools::install_github("Tsmnbx/ZooGVT")
library(ZooGVT)
```

## Components

Here we show the ubronIDS for the phenotype descriptions that we can find for pectoral fin in the zebra Fish

```{r}
IdDataFrame<-GetUberonIds("Danio_rerio","Zebrafish","pectoral fin")
IdDataFrame
```


To see what Species we can use for this package that are both in BgeeDB and have phenotype descriptions in  ensembl as both are required run ListSpecies()
```{r}

ListSpecies()
```


To see what phenotype descriptions from ensembl are available and what could be used for the feature parameter run ListPhenotypesDescriptions()
```{r}

ListPhenotypesDescriptions()
```


Lets do an example of a species
In the code below we ran GenesFromAnatomy to get a radar graph of the number of genes we get for the phenotype descriptions of a feature
```{r}

AnatomyGeneRadarGraph <-GenesFromAnatomy("Danio_rerio","Zebrafish","pectoral fin")
```

Here we ran the GeneAnatomyParty to get multiple graphs for the different species for the same feature. Its tricky you have to have an feature that is just precise enough to get good results
```{r eval=FALSE}

ListofAnimals<-c("Zebrafish","Human")
GeneAnatomyParty(ListofAnimals,"Cardiac")
```

## Package References

[Efe, J. (2020) ZooGVT: A R Package For
 BCB410H. Unpublished.](https://github.com/Tsmnbx/ZooGVT)

<br>

  ## Other References

  R Core Team (2020). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. https://www.R-project.org/

  Bastian F et al. (2008). Bgee: Integrating and Comparing Heterogeneous
Transcriptome Data Among Species. Data Integration Life Sci. Lecture Notes
in Computer Science, 5109, 124-31


Martin Morgan (2019). BiocManager: Access the Bioconductor Project Package
Repository. R package version 1.30.10.
https://CRAN.R-project.org/package=BiocManager

Mapping identifiers for the integration of genomic datasets with the
R/Bioconductor package biomaRt. Steffen Durinck, Paul T. Spellman, Ewan
Birney and Wolfgang Huber, Nature Protocols 4, 1184-1191 (2009).

BioMart and Bioconductor: a powerful link between biological databases
and microarray data analysis. Steffen Durinck, Yves Moreau, Arek
Kasprzyk, Sean Davis, Bart De Moor, Alvis Brazma and Wolfgang Huber,
Bioinformatics 21, 3439-3440 (2005).


Minato Nakazawa (2019). fmsb: Functions for Medical Statistics Book
with some Demographic Data. R package version 0.7.0.
https://CRAN.R-project.org/package=fmsb





